Stochastic Convolutional Sparse Coding

### 2018-11-08 cvpr2019/Results.tex
WH edits on full text

\section{Results} \label{sec:result}
We first validate the proposed algorithms on fruit and city datasets~\cite{zeiler2010deconvolutional}, both of which contain 10 training images in the size of $100 \times 100$. The online-mode algorithms are then adapted to one thousand $100 \times 100$ image patches randomly picked up from ImageNet~\cite{deng2009imagenet}. Note that batch-based CSC commonly can only handle less than 100 images simultaneously. The dictionary dimension size is set to $11 \times 11 \times 100$ in all experiments except for over-complete dictionary. All training and evaluation processes in this manuscript are performed on contrast normalized images~\cite{zeiler2010deconvolutional,heide2015fast}.

@@

\begin{subfigure}{0.23\textwidth}
  \includegraphics[width=1\linewidth]{figure/batchFruit100.pdf}
\end{subfigure}
\begin{subfigure}{0.23\textwidth}
  \includegraphics[width=1\linewidth]{figure/heideFruit100.pdf}
\end{subfigure}

\caption{Top: Convergence comparison between the-state-of-art method~\cite{heide2015fast} and the proposed method with different subsampling rate, all of which on performed on fruit dataset. Convergence is evaluated by monitoring the objective value of Eq.~\ref{eq:CSCmodel} on training images versus iterations and time, respectively. Bottom: Learned filters by the proposed method with $p=0.1$ and the comparable method. In these represented learned filters, our method learns more smooth Gabor-like filters and less number of noise-like filters with the same $\lambda$. Beyond this, it runs faster with regard to each iteration and shows better convergence behavior.}
\label{fig:subsampleResult}
\end{figure}

@@

{\bfseries Convergence.} Comparisons of the convergence between the proposed method (SBCSC) and the state-of-the-art batch-mode algorithm~\cite{heide2015fast} are shown in Fig.\ \ref{fig:subsampleResult}. A different selection of the subsampling rate reveals that the proposed strategy will slightly influence the convergence and the training objective of the minimization problem. Specifically, the more subsampled, the relatively slower convergence and the higher objective will be obtained. On the other hand, small subsampling rate will significantly accelerate the computation process, where $10\%$ subsampling achieves about $6 \times$ speedup over no subsampled solver and $2 \times$ speedup over state-of-the-art frequency solver for one iteration. We observe that a subsampling rate between $0.1$ and $0.2$ delivers empirically good enough convergence in our settings, as well as achieving at least $3 \times$ speedup. In general, the proposed method with various subsampling rates converges at around 10-12 iterations in all testing cases, acting similar to the competing methods. Notice that the comparable method has a relatively high objective at first $6-8$ iterations due to different splitting strategies are applied and subproblems are interleaved on the primary variables~\cite{wohlberg2016efficient}, though it converges to an optimum with comparable objective after 10 iterations. In summary, the convergence behaviors of the proposed algorithm is only slightly influenced by the subsampling strategy within the testing subsampling rates. Comparing to the state-of-the-art frequency solver, the proposed stochastic spatial-domain solver with a subsampling rate of $0.1$ reduces the computation time by a factor of two for the tested example. The robustness of the proposed algorithm is evaluated by additional experiments. Please see supplementary materials for reference.

{\bfseries Reconstruction}. The learned filters by SBCSC with subsampling rate $0.1$ are shown in the bottom of Fig.\ \ref{fig:subsampleResult}. For a visual comparison, we also show the filters learned from the competing method (bottom left of Figure 3 in~\cite{heide2015fast}). As can be observed, both of them learn some seemingly similar Gabor-like filters. While after zooming (we also show zoom-in comparisons in supplement), it can be seen that our learned Gabor-like filters have less impure structures. Moreover, the compared dictionary contains a number of noise-like filters, which are rarely existed in ours. We then demonstrate the effectiveness of the reconstructed filters in the application of image inpainting, which refers to reconstructing a full image from partial measurements. A numerical comparison of the reconstruction quality is shown in Fig.\ \ref{fig:PSNRrecon}. The filters learned by the proposed method not only show more visually decent structures and less specific features, but also demonstrate its capability to better reconstruct partial observed images with respect to the PSNR value. In addition to that, the learning process only takes about $50\%$ of the execution time than the comparison.

\subsection{Online Learning Settings}

{\bfseries Convergence.} Unlike the batch-based learning approaches which evaluate its convergence by monitoring the objective value on training datasets, a common way to evaluate the learning process of online learning model is to monitor its objective value on testing datasets. The testing objectives of the proposed method (SOCSC) and prior published online frequency-domain CSC method~\cite{liu-2018-first}, with respect to iterations, are shown in Fig.\ \ref{fig:onlineSmall}. In the same figure, we also keep track of the capability of the updated filters during the learning process to sparsely represent the testing images, which is demonstrated by the time evolution of PSNR. These two approaches stop at optimum positions with close testing objective values, though, they exhibit slightly different convergence behaviors with respect to iterations. The final testing PSNR values for both methods also reveal a similar reconstruction performance of the learned filters. As for the runtime comparison, the proposed method runs at least 5 times faster than the comparable method. Supplement shows additional comparisons for the other datasets.

@@

{\bfseries Over-complete dictionary.} As far as we know, none of the precedent CSC work reported or analyzed the over-complete dictionary (number of the dictionary is more than its degrees of freedom). One of the reasons could be that most of the prior work is batch-based, thus learning over-complete dictionary from small datasets would cause the overfitting issue, which may contain quite a few data-specific filters, and therefore be short of the generalization ability (we verify this explanation in supplement). The proposed online-based learning strategy (SOCSC) can overcome this issue by scaling the model up to arbitrary sample sizes.

@@

\begin{subfigure}{1\textwidth}
    \centering
  \includegraphics[width=0.75\linewidth]{figure/online100-filter.pdf}
\end{subfigure}
\end{minipage}
\begin{minipage}{0.6\textwidth}
\centering
\includegraphics[width=0.97\linewidth]{figure/online400-filter.pdf}
\vspace*{2mm}
\end{minipage}
\begin{minipage}{1\textwidth}
    \centering
    \resizebox{0.8\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR by (a) & 29.58 & 28.19 & 29.44 & 29.63 & 28.89 & 29.33 & 28.13  & 30.14 & 27.42 & 30.89 \\
            \hline
            PSNR by (b) & 29.63 & 28.22 & 29.57 & 29.90 & 29.12 & 29.59 & 28.05  & 30.17 & 27.53 & 31.08 \\
            \hline
            PSNR by (c) & \textbf{30.24} & \textbf{28.34} & \textbf{29.95}  & \textbf{30.30} & \textbf{29.43} & \textbf{29.96} & \textbf{28.24} & \textbf{30.57} & \textbf{27.72} & \textbf{31.67} \\
            \hline
        \end{tabular} }
\end{minipage}
\caption{Filters learned from 1000 images by our method (SOCSC) and the state-of-the-art online method~\cite{liu-2018-first}, and their corresponding reconstruction quality using the same hyperparameter settings as in Fig.\ \ref{fig:PSNRrecon}. (a) The under-complete dictionary ($11 \times 11 \times 100$) learned by~\cite{liu-2018-first}; (b) The under-complete dictionary ($11 \times 11 \times 100$) learned by SOCSC. (c) The over-complete dictionary ($11 \times 11 \times 400$) learned by SOCSC. These under-complete dictionaries, mainly composed of Gabor-like filters, can be seen as a subset of the represented over-complete dictionary, which contains a number of extra low contrast image features.}
\label{fig:overCompleteDic}
\end{figure*}

We demonstrate its capability on 1000 image patches with the size of $100 \times 100$, and learn $11 \times 11 \times 400$ over-complete dictionary, all of which are shown in Fig.\ \ref{fig:overCompleteDic} . For a visual comparison, 100 learned filters by the same algorithm and its comparable approach~\cite{liu-2018-first} are also represented. As can be observed, both of the approaches learn visually similar under-complete dictionary, while the proposed method takes $5 \times$ less training time than the competitive one. As for the learned over-complete dictionary, it is composed of the Gabor-like image features as represented in the under-complete dictionaries, as well as a number of low contrast features which are rarely existed in those under-complete dictionaries. This additional feature information would play an essential role to better reconstruct the natural images, which can be demonstrated by the results of reconstruction quality shown at the bottom of the same figure.

We further show a bottleneck revealed by the under-complete dictionary in the online-based CSC model. The top of Fig.\ \ref{fig:overComDicAndMinibatch} demonstrates that no more apparent progress could be observed when the number of training images is higher than a specific value for both of the online approaches ($K=100$). However, owing to more abundant filters, learning over-complete dictionary overcomes this bottleneck, and it shows a considerable improvement, with regard to PSNR, to have a sparse representation of the testing images during the learning process. This is also verified by applying these learned dictionaries to the image inpainting application as discussed above.

{\bfseries Mini-batching.} In practice, a mini-batching strategy would be preferred in order to gain advantages from modern parallel computing architectures. This is also a standard extension to stochastic optimization algorithms~\cite{Takac2013, PCDM, SCSG}. We denote the mini-batch size as $\eta$. In the proposed online algorithm (SOCSC), the time complexity for one step dictionary update will not increase linearly with the increase of $\eta$. Concretely speaking, updating $\code$ can be implemented by caching the Cholesky decomposition, and one computation of the matrix factorization can be applied to all of the currently selected batches. Herein, the complexity for doing updating $\code$ $\eta$ times all together is cheaper than $\eta$ times the complexity of updating one $\code$. In addition, the time complexity for updating $\filter$ is not linearly affected by the value of $\eta$, which will be executed only once in each training step regardless of $\eta$. One exception is to update surrogate matrices, the complexity of which is $\eta$ linearly dependent, while this step is not dominant in the runtime.

The runtime comparisons for various mini-batch sizes are shown in the bottom of Fig.\ \ref{fig:overComDicAndMinibatch}. Note that larger $\eta$ will result in less number of iterations to process all 1000 samples. The plots show that a larger mini-batch size will generally lead to a greater progress in first few training steps, though it takes additional running time and memory. Overall, mini-batched updates provides a more runtime efficient learning process in the online-based CSC model, and according to the obtained experimental results, $\eta=20$ achieves one magnitude speedup over $\eta=1$ to reach a condition with the same level convergence on a single-core program. Since computing sparse codes is a data-independent process, this mini-batched approach can be further accelerated in a distributed-computing system.
\begin{figure}[h]
\centering
\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/overComplete-ite.pdf}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/minibatch.pdf}
\end{subfigure}

@@

>>>>

\section{Results} \label{sec:result}
We first validate the proposed algorithms on the fruit and city
datasets~\cite{zeiler2010deconvolutional}, which each consist of 10
training images of size $100 \times 100$. The online-mode algorithms
are then adapted to one thousand $100 \times 100$ image patches
randomly picked up from ImageNet~\cite{deng2009imagenet}. Note that
batch-based CSC commonly can only handle less than 100 images
simultaneously. The dictionary size is set to $100$ filters of size
$11 \times 11$ pixels in all experiments except for over-complete
dictionary. All training and evaluation processes in this manuscript
are performed on contrast normalized
images~\cite{zeiler2010deconvolutional,heide2015fast}.

@@

\begin{subfigure}{0.23\textwidth}
  \includegraphics[width=1\linewidth]{figure/batchFruit100.pdf}
\end{subfigure}
\begin{subfigure}{0.23\textwidth}
  \includegraphics[width=1\linewidth]{figure/heideFruit100.pdf}
\end{subfigure}
\caption{Top: Convergence comparison between the-state-of-art method~\cite{heide2015fast} and the proposed method with different subsampling rate, all of which on performed on fruit dataset. Convergence is evaluated by monitoring the objective value of Eq.~\ref{eq:CSCmodel} on training images versus iterations and time, respectively. Bottom: Learned filters by the proposed method with $p=0.1$ and the comparable method. In these represented learned filters, our method learns more smooth Gabor-like filters and less number of noise-like filters with the same $\lambda$. Beyond this, it runs faster with regard to each iteration and shows better convergence behavior.}
\label{fig:subsampleResult}
\end{figure}

@@

{\bfseries Convergence.} Comparisons of the convergence between the
proposed method (SBCSC) and the state-of-the-art batch-mode
algorithm~\cite{heide2015fast} are shown in
Fig.~\ref{fig:subsampleResult}. A different selection of the
subsampling rate reveals that the proposed strategy will slightly
influence the convergence and the training objective of the
minimization problem. Specifically, the more subsampled, the
relatively slower convergence and the higher objective will be
obtained. On the other hand, small subsampling rate will significantly
accelerate the computation process, where $10\%$ subsampling achieves
about $6 \times$ speedup over the not subsampled spatial domain solver
and a $2 \times$ speedup over state-of-the-art Fourier-domain solver
for one iteration. We observe that a subsampling rate between $p=0.1$
and $p=0.2$ delivers empirically good enough convergence in our
settings, as well as achieving at least $3 \times$ speedup. In
general, the proposed method with various subsampling rates converges
at around 10-12 iterations in all testing cases, acting similar to the
competing methods.
\note{Not sure what you want to say here -- fix grammar:}
Notice that the comparison method has a relatively
high objective value during the first $6-8$ iterations due to
different splitting strategies are applied and subproblems are
interleaved on the primary variables~\cite{wohlberg2016efficient},
though it converges to an optimum with comparable objective after 10
iterations. In summary, the convergence behaviors of the proposed
algorithm is only slightly influenced by the subsampling strategy
within the testing subsampling rates. Comparing to the
state-of-the-art frequency solver, the proposed stochastic
spatial-domain solver with a subsampling rate of $0.1$ reduces the
computation time by a factor of two for the tested example. The
robustness of the proposed algorithm is evaluated by additional
experiments. Please see supplementary materials for reference.

{\bfseries Reconstruction}. The SBSC-learned filters with a
subsampling rate of $p=0.1$ are shown in the bottom of
Fig.~\ref{fig:subsampleResult}. For a visual comparison, we also show
the filters learned from the competing method (bottom left of Figure 3
in~\cite{heide2015fast}). As can be observed, both of them learn some
seemingly similar Gabor-like filters. However, zooming into the
details (please see supplemental materials for high resolution
versions) reveals that our filters appear less noisy than the ones
from~\cite{heide2015fast}. We then demonstrate the effectiveness of
the reconstructed filters in the application of image inpainting,
which refers to reconstructing a full image from partial
measurements. A numerical comparison of the reconstruction quality is
shown in Fig.~\ref{fig:PSNRrecon}. The filters learned by the proposed
method not only appear less noisy by themselves, but also achieve
better reconstruction quality on partially observed images with
respect to the PSNR value. Furthermore, the learning process
only takes about $50\%$ of the time for our method.\todo{specific
  times here!}.
\note{So, how does Bis' work from last year compare in terms of
  quality and time? It should
  have much better dictionaries than~\cite{heide2015fast} as well.}

\subsection{Online Learning}

{\bfseries Convergence.} Unlike the batch-based learning approaches
which evaluate its convergence by monitoring the objective value on
training datasets, a common way to evaluate the learning process of
online learning model is to monitor its objective value on test
datasets. In Fig.~\ref{fig:onlineSmall} we plot the objective values
against the iteration number for the proposed method (SOCSC) and a
recent online frequency-domain CSC method~\cite{liu-2018-first}. In
the same figure, we also keep track of the capability of the updated
filters during the learning process to sparsely represent the test
images, which is demonstrated by the time evolution of PSNR. These two
approaches stop at optimum positions with similar objective values.
%although they exhibit slightly different convergence behaviors with
%respect to iterations.
The final PSNR values for both methods also reveal a similar
reconstruction performance of the learned filters. In terms of runtime
comparison, however, the proposed method runs at least 5 times faster
than the comparison method. The supplement shows additional comparisons
for the other datasets.

@@

{\bfseries Over-complete dictionary.} To our knowledge, none of the
existing CSC work reported or analyzed the over-complete dictionary
(number of the dictionary is more than its degrees of freedom). One of
the reasons could be that most of the prior work is batch-based, thus
learning over-complete dictionary from small datasets would cause
overfitting issues, which may contain quite a few data-specific
filters, and therefore limit the ability to generalize the filters to
other data (we verify this explanation in the supplement). The
proposed online-based learning strategy (SOCSC) can overcome this
issue by scaling the model up to arbitrary sample sizes.

@@

\begin{subfigure}{1\textwidth}
    \centering
  \includegraphics[width=0.75\linewidth]{figure/online100-filter.pdf}
\end{subfigure}
\end{minipage}
\begin{minipage}{0.6\textwidth}
\centering
\includegraphics[width=0.97\linewidth]{figure/online400-filter.pdf}
\vspace*{2mm}
\end{minipage}
\begin{minipage}{1\textwidth}
    \centering
    \resizebox{0.8\linewidth}{!}{
        \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
            \cline{1-11}
            Image & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
            \hline
            PSNR by (a) & 29.58 & 28.19 & 29.44 & 29.63 & 28.89 & 29.33 & 28.13  & 30.14 & 27.42 & 30.89 \\
            \hline
            PSNR by (b) & 29.63 & 28.22 & 29.57 & 29.90 & 29.12 & 29.59 & 28.05  & 30.17 & 27.53 & 31.08 \\
            \hline
            PSNR by (c) & \textbf{30.24} & \textbf{28.34} & \textbf{29.95}  & \textbf{30.30} & \textbf{29.43} & \textbf{29.96} & \textbf{28.24} & \textbf{30.57} & \textbf{27.72} & \textbf{31.67} \\
            \hline
        \end{tabular} }
\end{minipage}
\caption{Filters learned from 1000 images by our method (SOCSC) and the state-of-the-art online method~\cite{liu-2018-first}, and their corresponding reconstruction quality using the same hyperparameter settings as in Fig.~\ref{fig:PSNRrecon}. (a) The under-complete dictionary ($11 \times 11 \times 100$) learned by~\cite{liu-2018-first}; (b) The under-complete dictionary ($11 \times 11 \times 100$) learned by SOCSC. (c) The over-complete dictionary ($11 \times 11 \times 400$) learned by SOCSC. These under-complete dictionaries, mainly composed of Gabor-like filters, can be seen as a subset of the represented over-complete dictionary, which contains a number of extra low contrast image features.}
\label{fig:overCompleteDic}
\end{figure*}

We demonstrate this ability on 1000 image patches with the size of
$100 \times 100$, and learn an $11 \times 11 \times 400$ over-complete
dictionary, which is shown in Fig.~\ref{fig:overCompleteDic}. For a
visual comparison, we also show $100$ learned filters generated by the
same algorithm and another 100 filters generated
by~\cite{liu-2018-first}. As can be observed, both of the approaches
learn visually similar under-complete dictionary, while the proposed
method takes $5 \times$ less training time than the comparison
method. The learned over-complete dictionary is composed of
the Gabor-like image features as represented in the under-complete
dictionaries, as well as a number of low contrast features which are
not typical for under-complete dictionaries. This additional
feature information would play an essential role to better reconstruct
the natural images, which can be demonstrated by the results of
reconstruction quality shown at the bottom of the same figure.

We further show a bottleneck revealed by the under-complete dictionary
in the online-based CSC model. The top of
Fig.~\ref{fig:overComDicAndMinibatch} demonstrates that no more
apparent progress could be observed when the number of training images
is higher than a specific value for both of the online approaches
($K=100$). However, owing to more abundant filters, learning
over-complete dictionary overcomes this bottleneck, and it shows a
considerable improvement in the PSNR of image representations.  This
is also verified by applying these learned dictionaries to the image
inpainting application as discussed above.

{\bfseries Mini-batching.} In practice, a mini-batching strategy would
be preferred in order to gain advantages from modern parallel
computing architectures. This is also a standard extension to
stochastic optimization algorithms~\cite{Takac2013, PCDM, SCSG}. We
denote the mini-batch size as $\eta$. In the proposed online algorithm
(SOCSC), the time complexity for one step dictionary update will not
increase linearly with the increase of $\eta$. Concretely speaking,
updating $\code$ can be implemented by caching the Cholesky
decomposition, and one computation of the matrix factorization can be
applied to all of the currently selected batches. Herein, the
complexity for doing updating $\code$ $\eta$ times all together is
cheaper than $\eta$ times the complexity of updating one $\code$. In
addition, the time complexity for updating $\filter$ is not linearly
affected by the value of $\eta$, which will be executed only once in
each training step regardless of $\eta$. One exception is the update
of surrogate matrices which has a complexity that is linear in
$\eta$. However, this step is not dominant in the runtime.

The runtime comparisons for various mini-batch sizes are shown in the
bottom of Fig.~\ref{fig:overComDicAndMinibatch}. Note that larger
$\eta$ will result in a smaller number of iterations to process all
$1000$ samples. The plots show that a larger mini-batch size will
generally lead to a greater progress in first few training steps,
though it takes additional running time and memory. Overall,
mini-batched updates provides a more runtime efficient learning
process in the online-based CSC model, and according to the obtained
experimental results, $\eta=20$ achieves a one order of magnitude
speedup over $\eta=1$ to reach a a comparable level of
convergence. Since computing sparse codes is a data-independent
process, this mini-batched approach can be further accelerated in a
distributed-computing system.
\begin{figure}[h]
\centering
\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/overComplete-ite.pdf}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \includegraphics[width=1\linewidth]{figure/minibatch.pdf}
\end{subfigure}

@@

% --- DO NOT DELETE ---
% Local Variables:
% mode: latex
% mode: flyspell
% mode: TeX-PDF
% End:
